---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r loading_data, echo=TRUE}
        data <- read.csv("activity.csv")
        head(data)
```


## What is mean total number of steps taken per day?

```{r mean_steps, echo=TRUE, warning=FALSE, message=FALSE}
        library(dplyr)
        mean_steps <- data %>% group_by(date) %>% 
                summarise(sum = sum(steps, na.rm= TRUE), mean = mean(steps, na.rm= TRUE))
        print(mean_steps, n = 62)
        hist(mean_steps$sum, col = 'red', xlab = 'Total Number of Steps', 
             main = 'Histogram of Total Number of Steps per Day' )
```

## What is the average daily activity pattern?

```{r, avg_daily_activity, echo = TRUE, warning = FALSE}
        daily_pattern <- data %>% group_by(interval) %>%
        summarise(mean = mean(steps, na.rm = TRUE))

        library(ggplot2)
        daily_pattern %>% ggplot(aes(interval, mean))+
                geom_line()+
                geom_segment(aes(x = daily_pattern$interval[which.max(daily_pattern$mean)], 
                                 xend = daily_pattern$interval[which.max(daily_pattern$mean)], 
                                 y=0, 
                                 yend=max(daily_pattern$mean)), 
                             col = 'red')
```

## Imputing missing values

``` {r, imputing, echo = TRUE, warning = FALSE, message = FALSE}
sum(is.na(data))

print(data %>% group_by(date) %>% summarise(sum = sum(is.na(steps))), n=62)
```


To impute these NA values we'll be use mean values of 5min intervals


``` {r, imputing_step2, echo = TRUE, message = FALSE, warning = FALSE}

        tidy <- data
        tidy <- left_join(tidy, daily_pattern, by = 'interval') %>%
        mutate(steps = ifelse(is.na(steps), mean, steps)) %>% 
                select(steps, date, interval)

```

```{r, total_num, echo = TRUE, message = FALSE, warning = FALSE}

        tidy_sum <- tidy %>% group_by(date) %>% summarise(sum = sum(steps))
        hist(tidy_sum$sum, col = 'red', xlab = 'Total Number of Steps', 
             main = 'Histogram of Total Number of Steps per Day')

```

Summary of Total Number of Steps per Day with without imputation;

``` {r, without, echo = TRUE, warning = FALSE, message = FALSE}

        data %>% group_by(date) %>% mutate(sum = sum(steps)) %>% 
        select(sum) %>% summary()

```
Summary of Total Number of Steps per Day with imputation;

``` {r, with, echo = TRUE, warning = FALSE, message = FALSE}

        tidy_sum %>% summary()

```

As seen in summaries there are no impact of imputation on the data. But getting rid of NA's is healthy way of running analyses.

## Are there differences in activity patterns between weekdays and weekends?

First of all date column must be converted date format to character format;

```{r, date, echo=TRUE, warning=FALSE, message=FALSE}

        tidy$date <- as.Date(tidy$date, "%Y-%m-%d")
```

Then add days column to data set;

```{r, days, echo=TRUE, warning=FALSE, message=FALSE}

        tidy <- tidy %>% mutate (days = weekdays(date))
```

Turn days column to factor column with respect to weekdays or weekends;

```{r, weekend, echo=TRUE, warning=FALSE, message=FALSE}

        tidy$days <- factor(tidy$days, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"), labels = c("Weekday", "Weekday", "Weekday", "Weekday", "Weekday", "Weekend", "Weekend"))

```

```{r, timeplot_weekdays, echo=TRUE, warning=FALSE, message=FALSE}

        library(gridExtra)


        graph_1 <- tidy %>% filter(days == "Weekday") %>% group_by(interval) %>% 
                summarise(mean = mean(steps)) %>% 
                ggplot(aes(interval, mean))+
                geom_line()+
                ggtitle("Weekdays")
                


        graph_2 <- tidy %>% filter(days == "Weekend") %>% group_by(interval) %>% 
                summarise(mean = mean(steps)) %>%
                ggplot(aes(interval, mean))+
                geom_line()+
                ggtitle("Weekends")
        
        grid.arrange(graph_1, graph_2, nrow = 2)

```



